\MOC Server Pages:\\Web Publishing as Staged Computation

### 2007-04-16 paper/paper.tex
svn-r30: added citation to Takebe
darcs-hash:20070416223508-e5a07-36329b3c836c46dff2cb183c1198a6bfb29f939c.gz

Strictly speaking, these languages are not exclusively confined to the
stages that I have indicated: Javascript can be run server-side, PHP
can be run off-line, and so on.  Nevertheless, migrating code between
stages is hard, and the need for quoting and persistence are
practically show-stoppers.  For comparison purposes (further described
in section~\ref{sec:perf}), we staged some code using PHP.  To achieve
persistence of composite variables from one stage to the next, it
contains gems like this:\label{php-gem}
\begin{verbatim}
    <?= "<? \$list = unserialize(\"".
        addcslashes(serialize($list),'"').
        "\"); ?>\n" ?>
\end{verbatim}
where the \texttt{serialize} library function occurs in stage one and
the \texttt{unserialize} in stage two.  Notice that the \verb+$+
preceding the first occurrence of \verb+list+ is quoted, but the
second occurrence is not.  The \texttt{addcslashes} function is needed
in case the serialized representation contains special characters
(such as double quotes or newlines) that would be misinterpreted by
the PHP parser in the next stage.  Interpreting the stage-one program
guarantees nothing about the well-formedness of the stage-two program
(generated as a string).

@@

@@

Sincere thanks to the anonymous reviewers, for many helpful and
enlightening recommendations on both the work and the presentation.
Thanks to Walid Taha for organizing the \MOC workshop, and to the rest
of the \MOC team for their development efforts.

>>>>

Strictly speaking, these languages are not exclusively confined to the
stages that I have indicated: Javascript can be run server-side, PHP
can be run off-line, and so on.  Nevertheless, migrating code between
stages is hard, and the need for quoting and persistence are
practically show-stoppers.  For comparison purposes (further described
in section~\ref{sec:perf}), we staged some code using PHP.  To achieve
persistence of composite variables from one stage to the next, it
contains gems like this:\label{php-gem}
\begin{verbatim}
    <?= "<? \$list = unserialize(\"".
        addcslashes(serialize($list),'"').
        "\"); ?>\n" ?>
\end{verbatim}
where the \texttt{serialize} library function occurs in stage one and
the \texttt{unserialize} in stage two.  Notice that the \verb+$+ %$
preceding the first occurrence of \verb+list+ is quoted, but the
second occurrence is not.  The \texttt{addcslashes} function is needed
in case the serialized representation contains special characters
(such as double quotes or newlines) that would be misinterpreted by
the PHP parser in the next stage.  Interpreting the stage-one program
guarantees nothing about the well-formedness of the stage-two program
(generated as a string).

@@

\citet{takebe02caching} developed a partial evaluation technique for
PHP.  Their tool, PHP-Mix, reads standard PHP source code and performs
binding time analysis.  Then, it generates a semi-static script as
output.  It understands a significant subset of the PHP function
library, including the database connectivity.  Unfortunately, some
properties of PHP---such as the lack of variable declarations and the
single global scope---cause the analysis to be overly conservative,
so it does not always achieve the optimizations we desire.

@@

Sincere thanks to the anonymous reviewers, for many helpful and
enlightening recommendations on both the work and the presentation.
%%
Thanks to my M.S. student, Kiichi Takeuchi, for exploring some
applications of these ideas, and for helping me understand the work of
\citet{takebe02caching} (written in Japanese).
%%
Thanks to Walid Taha for organizing the \MOC workshop, and to the rest
of the \MOC team for their development efforts.
